General constraints on magic state protocols from the statistical mechanics of \\Wigner negativity

### 2021-06-28 magic_paper.tex
midway through refs

% Figures
% Refs
% Notation
% Proofreading --- Citations
% Proofreading --- Main text
% Proofreading --- Appendices

% Figs check --- consistency (size, font, text size, pads, axes, cbars), captions, folder

@@

% Notation consistency --- noise vs error, \ln{} -> \ln, \ln vs \log vs \log_2, italics & \emph, Aberg -> {\AA}berg + [110 Knig], quasi-dist vs quasi-prob dist, def elbows, cumulant vs cumulative, stabilizer vs free, \p vs \w, W_\rho vs W_\rho(\bmz), \succ_r vs \succ
% UK vs US english

@@

This means in particular that $\E(\tau^{\otimes n})$ is a symmetric state on $m$ subsystems, so by the quantum de Finetti theorem~\cite{hudson_locally_1976, christandl_2007} we have that for $m \gg 1$
\begin{equation}
\E(\tau) \approx \int d\mu(x) \tau_x^{\prime \otimes m},
\end{equation}
where $d\mu(x)$ is a probability measure over a set $\{\tau'_x\}$ of single qudit states.

@@

%
%\begin{figure}
%    \centering
%    \includegraphics[scale=0.4]{figs/test/R_vs_E1.pdf}
%    \caption{Here, $(\epsilon, \epsilon', \beta) = (0.1, 0.0, 0.2)$, $H = \rm{diag}(0,E,10)$ and $H' = \rm{diag}(0,E',10)$.
%    }
%    \label{fig:rvse1}
%\end{figure}

%\begin{figure}
%    \centering
%    \includegraphics[scale=0.4]{figs/test/R_vs_E2.pdf}
%    \caption{Here, $(\epsilon, \epsilon', \beta) = (0.1, 0.0, 0.2)$, $H = \rm{diag}(0,5,E)$ and $H' = \rm{diag}(0,5,E')$.
%    }
%    \label{fig:rvse2}
%\end{figure}

%\begin{figure}
%    \centering
%    \includegraphics[scale=0.4]{figs/test/R_vs_H.pdf}
%    \caption{Here, $(\epsilon, \epsilon', \beta) = (0.1, 0.0, 0.2)$, $H = {\rm{diag}}(0,1,2)$ and $H' = (1-p-q){\rm{diag}}(0,1,2)+p{\rm{diag}}(1,2,0)+q{\rm{diag}}(2,0,1)$.
%    }
%    \label{fig:rvsh}
%\end{figure}

@@

However, it turns out that the analysis for the temperature-dependent bound given by Equation (\ref{eq:rate_bounds_proof}) implicitly made use of the single-shot R\'{e}nyi divergence $D_\infty (\p \hspace{1pt}||\hspace{1pt} \r)$, which is given by
\begin{equation}
	D_\infty (\p \hspace{1pt}||\hspace{1pt} \r) = \log \max_i \frac{p_i}{r_i}.
\end{equation}
It is clear that this does remain well-defined on quasi-distributions, and from the majorization relation also has a clear physical meaning, and in~\cref{app:lcst_technical}, we show that this divergence has the following properties that are derived via relative majorization.
\begin{theorem}\label{thm:Dinfty} 
	Let $\tau$ be in the interior of $\F$. Then $D_\infty(W_\rho || W_\tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(W_{\rho^{\otimes 2n}} \hspace{1pt}||\hspace{1pt} W_{\tau^{\otimes 2n}}) = n D_\infty(W_{\rho^{\otimes 2}} \hspace{1pt}||\hspace{1pt} W_{\tau^{\otimes 2}})$ for any $n \in \mathbb{N}$.
\item $D_\infty(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \geq D_\infty(W_{\E(\rho)} \hspace{1pt}||\hspace{1pt} W_{\E(\tau)})$ for any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$.
\end{enumerate}
\end{theorem}
The reason for the power of $2$ in the third property is that the maximization is sensitive to the presence of negativity in the Wigner distribution.

We can go further, and show that in fact a range of R\'{e}nyi entropies remain both well-defined and meaningful in the Wigner representation. The $\alpha$--R\'{e}nyi entropy $H_\alpha(\p)$ is defined as
\begin{equation}
	H_\alpha(\p) := \frac{1}{1-\alpha} \log \sum_i p_i^\alpha.
\end{equation}
To extend to quasi-distributions we must ensure the following: (a) the entropy is well-defined on quasi-distributions, and (b) obeys monotonicity under the majorization pre-order. In Appendix~\ref{app:single-shot} we prove the following.
\begin{theorem}\label{thm:HSchur} 
	If $\alpha = \frac{2a}{2b-1}$ for positive integers $a,b$ with $a \geq b$, then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'$ then $H_\alpha (\w) \leq H_\alpha(\w')$.
\end{theorem}
This monotonicity provides a meaningful extension of entropies to quasi-distributions. The unusual form of the $\alpha$ parameter is required so that the entropy function is well-defined and real-valued for all $\w$. The allowed values of $\alpha$ are dense in $\alpha >1$, however, while R\'{e}nyi entropies on probabilities can also be used for $0 < \alpha <1$ it turns out that for quasi-probability distributions this parameter regime is no longer monotonic in terms of the majorization pre-order (see Appendix~\ref{app:single-shot} for details).

@@

We also note that the R\'{e}nyi entropy can be described as a $q$--deformation of the Shannon entropy~\cite{baez2011renyi}, and that $H_\alpha(\p)$ can be related to the $q$--derivative of free energy.  For this, the R\'{e}nyi parameter becomes a temperature term $\alpha = \frac{T_0}{T}$ where $T_0$ is a reference temperature and derivatives are considered via the limit $T \rightarrow T_0$. Therefore, if we try to push this to the present setting, the presence of Wigner negativity would correspond to a divergence in the first derivative of an effective free energy as $T \rightarrow T_0$. It is interesting to speculate whether the presence of magic in Clifford circuits could admit a sensible interpretation in terms of zeroth-order phase transitions~\cite{Aharonov_2000, Shiyu_2020, Ippoliti_2021}.

\subsection{R\'{e}nyi divergences and general distillation bounds}
Every entropy can be obtained from a relative entropy~\cite{Gour_2020} and so we now state the extensions for R\'{e}nyi divergences of Wigner distributions.
The $\alpha$--R\'{e}nyi divergence $D_\alpha(\w||\r)$ is defined as
\begin{equation}
	D_\alpha(\w \hspace{1pt}||\hspace{1pt} \r) := \frac{1}{\alpha-1} \log \sum_i w_i^\alpha r_i^{1-\alpha},
\end{equation}
where $\w$ can be a quasi-distribution, $\r$ a probability distribution with positive components. In~\cref{app:single-shot}, we also show that the divergence satisfies the following.

@@

We can now apply this to obtain a family of magic distillation bounds based on R\'{e}nyi divergences.
\begin{theorem}\label{thm:entropic_bounds}
	Consider a general magic state distillation process on prime dimension qudits, that converts a magic state $\rho^{\otimes n} \longrightarrow \E(\rho^{\otimes n})=\rho'^{\otimes m}$and let $\tau$ be any full-rank stabilizer reference state on a qudit. Then the distillation rate $R := m/n$ is upper bounded as,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{\tilde{D}_\alpha( \rho', \tau')},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and the average divergence
	\begin{equation}
\tilde{D}_\alpha( \rho', \tau') := \frac{1}{m} D_\alpha (W_{\rho'^{\otimes m}} || W_{\tau'_m}),
\end{equation}
between the output magic state $\rho'^{\otimes m}$ and $\tau'_m = \E(\tau^{\otimes n})$.
\end{theorem}
The proof is given in Appendix~\ref{app:single-shot}.

While this bound is abstract in its present form, it does suggest future applications by viewing $\tilde{D}_\alpha( \rho', \tau')$ within the context of hypothesis testing, as a measure of distinguishability between $\rho'^{\otimes m}$ and the $m$ qudit state $\tau'_m$. 
However in order to properly link with hypothesis-testing one would have to first extend such results to quasi-distributions.

If correlations in $\tau'_m$ between subsystems can be neglected then $\tau'_m = \tau'^{\otimes m}$ for some qudit state $\tau'$ we have that
\begin{equation}
\tilde{D}_\alpha( \rho', \tau') = D_\alpha (W_{\rho'} || W_{\tau'}),
\end{equation}
which gives a generalized version of Theorem \ref{thm:free-energy}.

@@

We have described how relative majorization can be used to establish upper bounds on magic distillation protocols that take into account additional physics of the system. Our bounds exploited relatively simple aspects of the Lorenz curves of the quasi-distributions, so it would be of interest to sharpen these bounds and obtain a better handle on the Lorenz curve structure in the $n\rightarrow \infty$ limit. It would also be of interest to analyse what features of single-shot entropies can be extended to quasi-distributions in a sensible form [\ddd{cite}]. This raises interesting questions as we no longer have a notion of typicality and the central limit theorem does not apply. That said, for special states such as the Strange state, the asymptotic behaviour is relatively simple, so exact asymptotics for this should be possible. In~\cref{app:lcsu_technical} and~\cref{app:lcst_technical} we give additional results that may be of use for later work.

@@

\begingroup
\def\thetheorem{\ref{thm:entropic_bounds}}
\begin{theorem}
	Consider a general magic state distillation process that converts a magic state $\rho^{\otimes n} \longrightarrow \E(\rho^{\otimes n})=\rho'^{\otimes m}$and let $\tau$ be any full-rank stabilizer reference state. Then the distillation rate $R := m/n$ is upper bounded as,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{\tilde{D}_\alpha( \rho', \tau')},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ with the average divergence
	\begin{equation}
\tilde{D}_\alpha( \rho', \tau') := \frac{1}{m} D_\alpha (W_{\rho\prime_m} || W_{\tau'_m}),
\end{equation}
between the output magic state $\rho'^{\otimes m}$ and $\tau'_m = \E(\tau^{\otimes n})$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
	The bound is a direct consequence of the properties of the $\alpha$--R\'{e}nyi diveregence, proven in~\cref{thm:Da_props}.

@@

Since $\rho'^{\otimes m} \neq \tau'_m$, we have $D_\alpha(W_{\rho'_m} \hspace{1pt}||\hspace{1pt} W_{\tau'_m}) > 0$, which directly leads to the bound
\begin{equation}
	\frac{m}{n} \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{\tilde{D}_\alpha( \rho', \tau')}.
\end{equation}
\end{proof}

>>>>

% Aberg -> {\AA}berg
% Refs
% Notation
% Proofreading --- Citations
% Proofreading --- Main text
% Proofreading --- Appendices

%X Figs check --- consistency (size, font, text size, pads, axes, cbars), captions, folder

@@

% Notation consistency --- noise vs error, \ln{} -> \ln, \ln vs \log vs \log_2, italics & \emph, quasi-dist vs quasi-prob dist, def elbows, cumulant vs cumulative, stabilizer vs free, \p vs \w, W_\rho vs W_\rho(\bmz), \succ_r vs \succ
% UK vs US english

@@

This means in particular that $\E(\tau^{\otimes n})$ is a symmetric state on $m$ subsystems, so by the quantum de Finetti theorem~\cite{hudson_locally_1976, christandl_2007, renner_phd} we have that for $m \gg 1$
\begin{equation}
\E(\tau) \approx \int d\mu(x) \tau_x^{\prime \otimes m},
\end{equation}
where $d\mu(x)$ is a probability measure over a set $\{\tau'_x\}$ of single qudit states.

@@

@@

However, it turns out that the analysis for the temperature-dependent bound given by Equation (\ref{eq:rate_bounds_proof}) implicitly made use of the single-shot R\'{e}nyi divergence~\cite{renyi_1960} $D_\infty (\p \hspace{1pt}||\hspace{1pt} \r)$, which is given by
\begin{equation}
	D_\infty (\p \hspace{1pt}||\hspace{1pt} \r) = \log \max_i \frac{p_i}{r_i}.
\end{equation}
It is clear that this does remain well-defined on quasi-distributions, and from the majorization relation also has a clear physical meaning, and in~\cref{app:lcst_technical}, we show that this divergence has the following properties that are derived via relative majorization.
\begin{theorem}\label{thm:Dinfty} 
	Let $\tau$ be in the interior of $\F$. Then $D_\infty(W_\rho || W_\tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(W_{\rho^{\otimes 2n}} \hspace{1pt}||\hspace{1pt} W_{\tau^{\otimes 2n}}) = n D_\infty(W_{\rho^{\otimes 2}} \hspace{1pt}||\hspace{1pt} W_{\tau^{\otimes 2}})$ for any $n \in \mathbb{N}$.
\item $D_\infty(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \geq D_\infty(W_{\E(\rho)} \hspace{1pt}||\hspace{1pt} W_{\E(\tau)})$ for any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$.
\end{enumerate}
\end{theorem}
The reason for the power of $2$ in the third property is that the maximization is sensitive to the presence of negativity in the Wigner distribution.

We can go further, and show that in fact a range of R\'{e}nyi entropies remain both well-defined and meaningful in the Wigner representation. The $\alpha$--R\'{e}nyi entropy $H_\alpha(\p)$ is defined as~\cite{renyi_1960}
\begin{equation}
	H_\alpha(\p) := \frac{1}{1-\alpha} \log \sum_i p_i^\alpha.
\end{equation}
To extend to quasi-distributions we must ensure the following: (a) the entropy is well-defined on quasi-distributions, and (b) obeys monotonicity under the majorization pre-order. In Appendix~\ref{app:single-shot} we prove the following.
\begin{theorem}\label{thm:HSchur} 
	If $\alpha = \frac{2a}{2b-1}$ for positive integers $a,b$ with $a \geq b$, then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'$ then $H_\alpha (\w) \leq H_\alpha(\w')$.
\end{theorem}
This monotonicity provides a meaningful extension of entropies to quasi-distributions. The unusual form of the $\alpha$ parameter is required so that the entropy function is well-defined and real-valued for all $\w$. The allowed values of $\alpha$ are dense in $\alpha >1$, however, while R\'{e}nyi entropies on probabilities can also be used for $0 < \alpha <1$ it turns out that for quasi-probability distributions this parameter regime is no longer monotonic in terms of the majorization pre-order (see Appendix~\ref{app:single-shot} for details).

@@

We also note that the R\'{e}nyi entropy can be described as a $q$--deformation of the Shannon entropy~\cite{baez2011renyi}, and that $H_\alpha(\p)$ can be related to the $q$--derivative of free energy.  For this, the R\'{e}nyi parameter becomes a temperature term $\alpha = \frac{T_0}{T}$ where $T_0$ is a reference temperature and derivatives are considered via the limit $T \rightarrow T_0$. Therefore, if we try to push this to the present setting, the presence of Wigner negativity would correspond to a divergence in the first derivative of an effective free energy as $T \rightarrow T_0$. It is interesting to speculate whether the presence of magic in Clifford circuits could admit a sensible interpretation in terms of phase transitions~\cite{domb2000phase}.

\subsection{R\'{e}nyi divergences and general distillation bounds}
Every entropy can be obtained from a relative entropy~\cite{Gour_2020} so we now state the extensions for R\'{e}nyi divergences of Wigner distributions.
The $\alpha$--R\'{e}nyi divergence $D_\alpha(\w||\r)$ is defined as
\begin{equation}
	D_\alpha(\w \hspace{1pt}||\hspace{1pt} \r) := \frac{1}{\alpha-1} \log \sum_i w_i^\alpha r_i^{1-\alpha},
\end{equation}
where $\w$ can be a quasi-distribution, $\r$ a probability distribution with positive components. In~\cref{app:single-shot}, we also show that the divergence satisfies the following.

@@

We can now apply this to obtain a family of magic distillation bounds based on R\'{e}nyi divergences.
\begin{theorem}\label{thm:entropic_bounds}
	Consider a general magic state distillation process on prime dimension qudits, that converts a magic state $\rho^{\otimes n} \longrightarrow \E(\rho^{\otimes n})=\rho'^{\otimes m}$and let $\tau$ be any full-rank stabilizer reference state on a qudit. Then the distillation rate $R := m/n$ is upper bounded as,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{\rlap{\raisebox{-0.1ex}{$\widetilde{\phantom{D}}$}}D_\alpha( \rho', \tau')}
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and the average divergence
	\begin{equation}
\widetilde{D}_\alpha( \rho', \tau') := \frac{1}{m} D_\alpha (W_{\rho'^{\otimes m}} \hspace{1pt}||\hspace{1pt} W_{\tau'_m}),
\end{equation}
between the output magic state $\rho'^{\otimes m}$ and $\tau'_m = \E(\tau^{\otimes n})$.
\end{theorem}
The proof is given in Appendix~\ref{app:single-shot}.

While this bound is abstract in its present form, it does suggest future applications by viewing $\widetilde{D}_\alpha( \rho', \tau')$ within the context of hypothesis testing~\cite{tomamichel2013}, as a measure of distinguishability between $\rho'^{\otimes m}$ and the $m$ qudit state $\tau'_m$. 
However in order to properly link with hypothesis-testing one would have to first extend such results to quasi-distributions.

If correlations in $\tau'_m$ between subsystems can be neglected then $\tau'_m = \tau'^{\otimes m}$ for some qudit state $\tau'$ we have that
\begin{equation}
	\widetilde{D}_\alpha( \rho', \tau') = D_\alpha (W_{\rho'} \hspace{1pt}||\hspace{1pt} W_{\tau'}),
\end{equation}
which gives a generalized version of Theorem \ref{thm:free-energy}.

@@

We have described how relative majorization can be used to establish upper bounds on magic distillation protocols that take into account additional physics of the system. Our bounds exploited relatively simple aspects of the Lorenz curves of the quasi-distributions, so it would be of interest to sharpen these bounds and obtain a better handle on the Lorenz curve structure in the $n\rightarrow \infty$ limit. It would also be of interest to analyse what features of single-shot entropies can be extended to quasi-distributions in a sensible form~\cite{renyi_1960, renner_phd, tomamichel2013}. This raises interesting questions as we no longer have a notion of typicality and the central limit theorem does not apply. That said, for special states such as the Strange state, the asymptotic behaviour is relatively simple, so exact asymptotics for this should be possible. In~\cref{app:lcsu_technical} and~\cref{app:lcst_technical} we give additional results that may be of use for later work.

@@

\begingroup
\def\thetheorem{\ref{thm:entropic_bounds}}
\begin{theorem}
	Consider a general magic state distillation process on prime dimension qudits, that converts a magic state $\rho^{\otimes n} \longrightarrow \E(\rho^{\otimes n})=\rho'^{\otimes m}$and let $\tau$ be any full-rank stabilizer reference state on a qudit. Then the distillation rate $R := m/n$ is upper bounded as,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{\rlap{\raisebox{-0.1ex}{$\widetilde{\phantom{D}}$}}D_\alpha( \rho', \tau')}
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and the average divergence
	\begin{equation}
\widetilde{D}_\alpha( \rho', \tau') := \frac{1}{m} D_\alpha (W_{\rho'^{\otimes m}} \hspace{1pt}||\hspace{1pt} W_{\tau'_m}),
\end{equation}
between the output magic state $\rho'^{\otimes m}$ and $\tau'_m = \E(\tau^{\otimes n})$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
	The bound is a direct consequence of the properties of the $\alpha$--R\'{e}nyi diveregence, proven in~\cref{thm:Da_props}.

@@

Since $\rho'^{\otimes m} \neq \tau'_m$, we have $D_\alpha(W_{\rho'_m} \hspace{1pt}||\hspace{1pt} W_{\tau'_m}) > 0$, which directly leads to the bound
\begin{equation}
	\frac{m}{n} \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{\widetilde{D}_\alpha( \rho', \tau')}.
\end{equation}
\end{proof}
